The number of lattices of Voronoi's first kind

### 2014-06-06 findsuperbase/findsuperbase.tex
added neew build and ran new simulations.  This code probably doesn't work for special lattices with a combination of strict and lax relevant vectors

The plot shows the estimated (after 500 trials) probability of obtaining a lattice of Voronoi's first kind for $n=1,2,3,4$.  Three different methods for generating the $n \times n$ basis matrix $B$ are considered: 
\begin{enumerate}
\item entries drawn independently and uniformly in the interval $[0,1]$, 
\item entries drawn independently and Gaussian with zero mean and variance one, and 
\item the $i,j$th entry satisfies
\[
B_{i,j} = \begin{cases}
R_{i,j} & 1 \leq i,j \leq h \\
-C_{i-h,j} & h < i \leq n, 1 \leq j \leq h \\
C_{i,j-h} & h < j \leq n, 1 \leq i \leq h \\
R_{i-h,j-h} & h < i,j \leq n,
\end{cases}
\]
where $h = \ceil{n/2}$ is the smallest integer greater than or equal to $n/2$ and $R_{i,j}$ and $C_{i,j}$ for $i,j = 1,\dots,h$ are independently Gaussian distributed with zero mean and variance one. 
\end{enumerate}
When $n$ is even, the third case corresponds with the model commonly used for MIMO channels, that is, the basis matrix is of the form
\[
B = \left(\begin{array}{cc}
R & -C \\
C & R
\end{array}\right)
\]
where $R$ and $C$ are $\tfrac{n}{2} \times \tfrac{n}{2}$ matrices containing the real and imaginary parts of the channel coefficients.

As expected, all lattices generated in dimensions 1, 2, and 3 were found to be of Voronoi's first kind.  When $n=4$ roughtly 25\% of the lattices generated by methods 1 and 2 were of Voronoi's first kind, but none of the 500 bases generated by method 3 were of first kind.  I find this very interesting! The simulations suggest that the lattices of Voronoi's first kind do not lie within a space of measure zero with respect to the measures induced by the first two generation methods, but that they do lie on a space of measure zero with respect to the lattices that occur when converting complex valued MIMO channel matrices into real valued lattices bases.  This might suggest that lattices of first kind will (atleast not directly) be useful for MIMO.

I was too oppimistic when I thought I would be able to run this with $n=6$.  This would take too long with the existing code.  There are a number of optimisations that can be made that might make running simulations with $n=6$ possible.

\begin{figure}[tp]
\centering 
  \begin{tikzpicture}
    \begin{axis}[font=\footnotesize,clip mode=individual,height=7cm,width=9cm,legend style={draw=none,fill=none,legend pos=south west,font=\footnotesize}]    
      \addplot[mark=*,color=red,only marks,/pgfplots/error bars/.cd,
        x dir=none,
        y dir=both,
        y explicit] table[x=n,y=proportion,y error=stddev] {code/data/uniform};
      \addplot[mark=square,color=blue,only marks,/pgfplots/error bars/.cd,
        x dir=none,
        y dir=both,
        y explicit] table[x=n,y=proportion,y error=stddev] {code/data/gaussian};
      \addplot[mark=x,color=green,only marks,/pgfplots/error bars/.cd,
        x dir=none,
        y dir=both,
        y explicit] table[x=n,y=proportion,y error=stddev] {code/data/mimo};
\legend{Uniform, Gaussian, MIMO}
    \end{axis}
  \end{tikzpicture}
\caption{Estimated probability of generating a lattices of Voronoi's first kind for different basis distributions and dimension $n=1,2,3,4$.}
\end{figure}

>>>>

The plot shows the estimated (after 5000 trials) probability of obtaining a lattice of Voronoi's first kind for $n=1,2,3,4,5,6$.  Three different methods for generating the $n \times n$ basis matrix $B$ are considered: 
\begin{enumerate}
\item entries drawn independently and uniformly in the interval $[0,1]$, 
\item entries drawn independently and Gaussian with zero mean and variance one, and 
\item the $i,j$th entry satisfies
\[
B_{i,j} = \begin{cases}
R_{i,j} & 1 \leq i,j \leq h \\
-C_{i-h,j} & h < i \leq n, 1 \leq j \leq h \\
C_{i,j-h} & h < j \leq n, 1 \leq i \leq h \\
R_{i-h,j-h} & h < i,j \leq n,
\end{cases}
\] 
where $h = \ceil{n/2}$ is the smallest integer greater than or equal to $n/2$ and $R_{i,j}$ and $C_{i,j}$ for $i,j = 1,\dots,h$ are independently Gaussian distributed with zero mean and variance one. 
\end{enumerate}
When $n$ is even, the third case corresponds with the model commonly used for MIMO channels, that is, the basis matrix is of the form
\[
B = \left(\begin{array}{cc}
R & -C \\
C & R
\end{array}\right)
\]
where $R$ and $C$ are $\tfrac{n}{2} \times \tfrac{n}{2}$ matrices containing the real and imaginary parts of the channel coefficients.

As expected, all lattices generated in dimensions 1, 2, and 3 were found to be of Voronoi's first kind.  When $n=4$ roughly 25\% of the lattices generated by methods 1 and 2 were of Voronoi's first kind, but none of the 5000 bases generated by method 3 were of first kind for $n=4$.  When $n=5$ approximately 3\% of the lattices generated were of Voronoi's first kind for all 3 generation methods.  When $n=6$ less than 1\% of the lattices generated by methods 1 and 2 where of first kind and approximately 2\% of the lattices generated by methods 3 where of first kind.

It is interesting that none of the lattices corresponding with MIMO channel matrices are of first kind in dimension 4, but there are some in dimension 6.

%I find this very interesting! The simulations suggest that the lattices of Voronoi's first kind do not lie within a space of measure zero with respect to the measures induced by the first two generation methods, but that they do lie on a space of measure zero with respect to the lattices that occur when converting complex valued MIMO channel matrices into real valued lattice bases.  This might suggest that lattices of first kind will not be useful for MIMO (at least not directly).

%It would be interesting to see if these conclusion hold in higher dimension. I was too optimistic when I thought I would be able to run this with $n=6$.  This would take too long with the existing code.  There are a number of optimisations that can be made that might make running simulations with $n=6$ possible.

\begin{figure}[tp]
\centering 
  \begin{tikzpicture}
    \begin{axis}[ymode=log,font=\footnotesize,clip mode=individual,height=7cm,width=9cm,legend style={draw=none,fill=none,legend pos=south west,font=\footnotesize}]    
      \addplot[mark=*,color=red,only marks,/pgfplots/error bars/.cd,
        x dir=none,
        y dir=both,
        y explicit] table[x=n,y=proportion,y error=stddev] {code/data/uniform};
      \addplot[mark=square,color=blue,only marks,/pgfplots/error bars/.cd,
        x dir=none,
        y dir=both,
        y explicit] table[x=n,y=proportion,y error=stddev] {code/data/gaussian};
      \addplot[mark=x,color=green,only marks,/pgfplots/error bars/.cd, 
        x dir=none,
        y dir=both,
        y explicit] table[x=n,y=proportion,y error=stddev] {code/data/mimo};
\legend{Uniform, Gaussian, MIMO}
    \end{axis}
  \end{tikzpicture}
\caption{Estimated probability of generating a lattices of Voronoi's first kind for different basis distributions and dimension $n=1,2,3,4$.}
\end{figure}
